---
title: "Sheffield Weather Trends"
author: "Hazel C Gordon"
date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)

library(tidyverse)
library(here)
library(gganimate)

```

## Project Description

Todays class we will be looking at Sheffields weather trends. 

Data is available from this link:

https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/sheffielddata.txt 

The URL for these pages is:

https://hazelcgordon.github.io/sheffield_weather/

The repository for these pages is:

https://github.com/hazelcgordon/sheffield_weather.git

## Import Data

```{r import}

loc = "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/sheffielddata.txt"

# split values by 'tabs' with \t (white space)
# this approach doesn't work

#df <- read.csv(loc, sep = "\t", skip = 7)

# import as text, instead of csv, and clean 

dat <- readLines(loc)

# change the word 'provisional' to nothing, replace every instance 

dat <- gsub("Provisional", "", dat)
dat <- gsub("\\*", "", dat)
dat <- gsub("\\#", "", dat)

# save new file

writeLines(dat, here("data", "cleaned.txt"))

# import data file

df <- read.table("data/cleaned.txt", skip = 7, fill = TRUE)

```

## Wrangle Data

```{r wrangle}

cols <- c("year", "month", "max_t", "min_t", "af", "rain", "sun")

colnames(df) <- cols

```


## Visualise the Data

```{r plot}

# I need max temperature to be a number, use as.numeric

df$max_t <- as.numeric(df$max_t)

# create plot, change transparency with alpha

# canvas

p <- ggplot(df, aes(x = month, y = max_t, group = year, colour = year))

# ink

p + geom_line(alpha = 0.1) +
  geom_line(data = df %>% filter(year == 2024), colour = "red", linewidth = 2)

```

## Animated plot

We are going to use gganimate to create an animated plot. Think carefully about if the animation is distracting if considering this kind of plot.

```{r animated}

# create an average temp for each year

df2 <- df %>%
  group_by(year) %>%
  summarise(avg_t = mean(max_t))

# create an animated plot

p2 <- ggplot(df2, aes(x = year, y = avg_t), frame = year)

# cumulative keeps it on the screen

anim <- p2 + 
  geom_point() +
  transition_manual(year, cumulative = TRUE) +
  ggtitle("The year is: {frame+1882}")

anim

# export as gif, then save the plot to be embedded into github pages

```

